As we previously mentioned, NHQM arises when one observes the binding momentum, related to the energy eigenvalues as
\begin{equation}
    p=\sqrt{2mE}
\end{equatoin}
and the energy $E$ is complex.
This energy will be complex if we expand the hamiltonian in a Berggren-basis to describe resonance.

\section{The Berggren basis}
Previously, we calculated the energy eigenvalues in momentum-space \cref{sec:mom-space}, now we will introduce a generalization of this.
If we look at our momentum-space expansion, we see that it consists of some discrete momentum-values along the real axis.
This is fine as long as the states we observe are bound.
If they on the other hand are quasi-bound resonance-states, they will appear as poles in the complex momentum-plane.
With knowledge of Cauchy's Residue theorem, we know that we must integrate around this pole to take care of it's contribution.

The way to do this is to instead of using a strictly real basis introduce a complex one, called a Berggren-basis (Should we cite here?).
The Berggren-basis is expressed as a contour in the complex plane, the one we used, with slight modifications, is seen in \cref{fig:}.
By lowering this contour to contain the pole we include the resonance in the calculations and are thus able to predict its behavior.
Last one have to include the resonance pole in the basis for it to be complete\cite{berggren}.

{\Large Should this above move to introduction, or should we have it here?}

\section{Results}
By remodel our old mom-space solution to allow different contours, we were ready to get som results to see if our new model was correct.
{\Large RESULTS!!!!!!!!!!!!!!!}